---
title: "ACS Census Estimate Data: Process Data""
output: 
  html_notebook: default
  toc: TRUE
  toc_float: TRUE
---

# Note
  - This code proceses the ACS Census data from the socialexplorer.com website

<br> </br>

# Load workspace
```{r}
loadhistory("ACS_2015_5yr_MSA_Process_Data.Rhistory")
load("ACS_2015_5yr_MSA_Process_Data.RData")
```
<br> </br>

# Saving code progress criteria

```{r}

savehistory("ACS_2015_5yr_MSA_Process_Data.Rhistory")
save.image("ACS_2015_5yr_MSA_Process_Data.RData")
# Set working directory where data is located:
setwd("C:/Users/marqu/Documents/GitHub/Economics/ACS_SocialExplorer/ACS_2015_MSA_Estimate")
```
<br> </br>

# Import Packages
```{r}
source('~/R/R_startup.R' , encoding = 'UTF-8')
cat('\14')
sessionInfo()
```
<br></br>

# Input parameters needed for data processing

```{r}
#input_file <- 'ACS_2015_5yr_MSA_data.csv' # input for package
input_file <- 'C:\\Users\\marqu\\Documents\\GitHub\\Economics\\ACS_SocialExplorer\\ACS_2015_BlockGroup_Estimate\\ACS_2015_5yr_BlockGroup_data.csv'
output_Meta_data <- 'ACS_MSA_5yr_2015_Meta_Data.csv' # 'ACS_BG_5yr_2015_Meta_Data.csv'
output_ACS_data <- 'ACS_MSA_5yr_2015.csv' # 'ACS_BG_5yr_2015.csv'
output_ACS_data_tvs <- 'ACS_MSA_5yr_2015.tsv' # 'ACS_BG_5yr_2015.tsv'
output_ACS_data_bz2 <- 'ACS_MSA_5yr_2015.csv.bzip2' # 'ACS_BG_5yr_2015.csv.bzip2'
```
<br></br>

# Grab the first two rows that have the Variable name and the Variable description
```{r}
library(tcltk)


col_desc <- read_csv(input_file, n_max = 2, col_names = FALSE)
Variable_Name <- gather(col_desc[2,], value = "Variable_Name")
Variable_Description <- gather(col_desc[1,], value = "Variable_Description")
Meta_data <- left_join(Variable_Name, Variable_Description, by = "key")
Meta_data$key <- NULL
write_csv(Meta_data, output_Meta_data ) # input for package

```
<br></br>

# Import ACS MSA file
```{r}
# hadley read
ACS_Data <- read_csv(input_file, skip = 1 , n_max = 1000)
```
<br></br>


# Remove columns with NA in all rows
```{r}

ACS_Data <- ACS_Data[, !(colSums(is.na(ACS_Data)) == nrow(ACS_Data))]
```
<br></br>


# Convert Numeric data to Numeric
```{r}

ACS_Data[, 12:ncol(ACS_Data)] <- sapply(ACS_Data[, 12:ncol(ACS_Data)], as.numeric) # 12 msa 13/14 bg
```
<br></br>

# Put single quotes around text fields for HIVE Upload
```{r}

ACS_Data[, 2:3] <- sapply(ACS_Data[, 2:3], function(y) gsub(", " , "_",y) ) # 12 msa 13/14 bg
ACS_Data[, 2:3] <- sapply(ACS_Data[, 2:3], function(y) gsub("; " , "_",y) )
```

# Write the data to disk
```{r}
write_csv(ACS_Data , output_ACS_data )
#write_tsv(ACS_Data , output_ACS_data ) # Write as Tab-delimited for HIVE upload

# Compressing
#msa_zipped <- bzfile(output_ACS_data_bz2)
#write.csv(ACS_Data, bzfile(output_ACS_data_bz2))
```
<br></br>

# Clear the Environment
```{r}
rm(list = c(ls()))
```




